---
title: "Predator thermal stress - Thermal parameter derivation and analysis"
output: html_document
date: '2023-12-05'
---
# Install and load necessary packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # will automatically include code in the ouput unless we state otherwise
knitr::opts_chunk$set(warning= FALSE) # will NOT include warning messages in the output unless we state otherwise
# knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE) #keeps code wrapped in the output
# if you haven't installed these packages run the following line (first delete the # sign):

install.packages(c("tidyverse", "janitor", "kableExtra", "cowplot", "skimr", "measurements"))
library(tidyverse) #data wrangling & cleaning incl. ggplot
library(janitor) # more data cleaning functions
library(kableExtra) # making pretty tables
library(cowplot) # making publication quality figures or multi-panel figures
library(skimr) # tool to get quick overview of dataset
library(measurements) # useful to convert units
library(magrittr) # magrittr::
library(boot)
library(mime)
library(checkmate)
library(curl)
library(nls.multstart)
library(broom) #augment
library(boot)
library(nlstools)
library(proto)
library(nls2)
library(here)
library(lme4)
library(Matrix)
library(lmerTest)
library(stats)
library(Rmisc)
library(purrr) #map
library(Hmisc)
library(FSA)
library(rcompanion)
library(car)
library(multcompView)
library(lsmeans)
library(sjstats)
library(pwr)
library(jtools)
library(ggpubr)
library(multcomp)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggeffects)
library(wesanderson)
library(patchwork)
library(dplyr)

```

```{r}
#load the previously generated file containing the respiration rates for all individuals across temperatures

setwd("~/Desktop/TPC Project")

mydata<-read.csv('Photo.R_rates.csv')

#mydata$X<-NULL
View(mydata)

```

# Load the Sharpe-Schoolfield equation for high-temperature inactivation
```{r}
# define the Sharpe-Schoolfield equation
schoolfield_high <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) #units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)
}

```

# Fitting the model using the Sharpe-Schoolfield equation
```{r}
# fit over each set of groupings
#this takes 2-3 mins to run

mydata$K<-mydata$Temp.C + 273.15 # creating a column of  Temperature in Kelvins

fits <- mydata %>%
  group_by(., Organism.ID, Species, Experiment, fed.fasted, subtidal.intertidal, Season, predator.prey) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(ln.rate ~ schoolfield_high(lnc, E, Eh, Th, temp = K, Tc = TC),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = -5, Eh = -5, Th = 180),
                                                start_upper = c(lnc = 10, E = 5, Eh = 70, Th = 390),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = -10, Eh = -10, Th = -10))))

```

# Check out model fits
```{r}
# look at a single fit
summary(fits$fit[[1]])

# look at output object
#select(fits, ID, data, fit)  

# get summary info
info <- fits %>%
  unnest_legacy(fit %>% map(glance))

# get params
params <- fits %>%
  unnest_legacy(fit %>% map(tidy))
```

# Merging datafiles
```{r}
#left join params with meta data file to have treatment in data frame

params <- left_join(params, mydata, by = "Organism.ID")

View(params)
write.csv(params, "../Cni_phys_TPC/Output/params_table.csv")
```

# Estimating Confidence Intervals 
```{r}
#Estimating the confidence interval for the model
CI <- fits %>% 
  unnest_legacy(fit %>% map(~ confint2(.x) %>%
                       data.frame() #%>%
                  #     rename(., conf.low = X2.5.., conf.high = X97.5.. )
                     )) %>%
  dplyr::group_by(., Organism.ID) %>%
  dplyr::mutate(., term = c('lnc', 'E', 'Eh', 'Th')) %>%
  ungroup()

CI
```

# Changing names for CI
```{r}
colnames(CI)[7:8]<-c("conf.low", "conf.high") # rename columns
head(CI)

# merge parameters and CI estimates
params <- merge(params, CI, by = intersect(names(params), names(CI)))
head(params)

write.csv(params, "params_table.csv") #adding the CI to the parameter metadata matrix 

# get predictions
preds <- fits %>%
  unnest_legacy(fit %>% map(augment))


# new data frame of predictions, do this to set a sequence to make a smooth curve with your prediction points. We are making the prediction for a total 150 from the lowest temperature to the highest one. 

 new_preds <- mydata %>%  do(., data.frame(K = seq(min(.$K), max(.$K), length.out = 300), stringsAsFactors = FALSE)) #setting a specific sequence so you can have a smooth curve
 
 
max_min <- mydata %>% group_by(Organism.ID) %>%
  dplyr::summarise(., min_K = min(K), max_K = max(K)) %>%
  ungroup() 

```

# Generate new predictions
```{r}
preds2 <- fits %>%
  unnest_legacy(fit %>% map(augment, newdata = new_preds)) %>%
  merge(., max_min, by = "Organism.ID") %>%
  group_by(., Organism.ID) %>%
  filter(., K > unique(min_K) & K < unique(max_K)) %>%
  dplyr::rename(., ln.rate = .fitted) %>%
  ungroup()

write.csv(preds2, "preds2.csv")
```

```{r}
#left join preds2 with meta data file to have treatment on dataframe
preds2 <- left_join(preds2, mydata)
```


# Calculating Topt
```{r}
get_topt <- function(E, Th, Eh){
  return((Eh*Th)/(Eh + (8.62e-05 *Th*log((Eh/E) - 1))))
}

# Synthesizing data from all of the temperatures to one. I am summarizing (mean) all the rows for each species (110-12 temperatures). The mean of 10-12 values that are the same as the value (ie duplicates) 
params <-read.csv("params_table.csv")
params_short <- params %>%
  dplyr::select(Species, Organism.ID, term, estimate, subtidal.intertidal, Season, TC) %>%
  dplyr::group_by(Species, Organism.ID, term, subtidal.intertidal, Season, TC) %>%
  dplyr::summarise(estimate = mean(estimate))

View(params_short)

#Nyssa's code for Topt using the summarize tool on each  
Topt_data <- params_short %>%
  dplyr::select(Organism.ID, Species, subtidal.intertidal, Season, term, estimate, TC) %>%
  spread(term, estimate) %>%
  mutate(Topt = get_topt(E, Th, Eh)) %>%
  group_by(., Species, Organism.ID, subtidal.intertidal, Season, TC)

View(Topt_data)

#Converting topt to celsius 
Topt_data$Topt <- Topt_data$Topt - 273.15 
View(Topt_data)

write.csv(Topt_data, "Topt_parameters.csv")


Topt_summarise<-Topt_data %>%
  group_by(., Species, subtidal.intertidal, Season) %>%
  summarise_all(mean, na.rm=TRUE)

#Topt_summarise$Topt_C<-Topt_summarise$Topt-273.15

View(Topt_summarise)

write.csv(Topt_summarise, "Summary_Tp_parameters.csv")

```

# Derive Pmax parameter
```{r}
Pmax <- Topt_data %>%
  #dplyr::select(Species, Organism.ID, E, Eh, lnc, Th) %>%
  #group_by(Species, Organism.ID) %>%
  #summarise(Pmax = schoolfield_high(lnc = lnc, E = E, Th = Th , Eh = Eh, temp = Topt_data$Topt, Tc = 23))
  mutate(Pmax = schoolfield_high(lnc = lnc, E = E, Th = Th, Eh = Eh, temp = Topt + 273.15, Tc = TC))
 
write.csv(Pmax, "Pmax.csv")
```


# Plotting thermal performance curves (TPCs) by species across temperatures (Fig. 4A)
```{r}
library(ggplot2)
library(dplyr)

# Read data
preds2 <- read.csv('preds2_resp.csv')
mydata <- read.csv('Resp_preds.csv')

# Define the list of species you want to filter
species_list <- unique(mydata$Species)

# Filter data
mydata_filtered <- mydata %>% filter(Species %in% species_list)
preds2_filtered <- preds2 %>% filter(Species %in% species_list, predator.prey == "Predator")

# Define color and fill palette
color_palette <- c("Tplanospira" = "orangered1", 
                    "Hprinceps" = "orange1", 
                    "Vmelones" = "coral4", 
                    "Hcumingi" = "dodgerblue1")

# Define custom labels for the legend with italicized text
custom_labels <- c("Tplanospira" = expression(italic("T. planospira")),
                    "Hprinceps" = expression(italic("H. princeps")),
                    "Vmelones" = expression(italic("V. melones")),
                    "Hcumingi" = expression(italic("H. cumingi")))

# Manual computation of confidence intervals
compute_ci <- function(df) {
  fit <- loess(ln.rate ~ K, data = df, span = 0.2)
  pred <- predict(fit, newdata = df, se = TRUE)
  df$fit <- pred$fit
  df$lower <- pred$fit - 1.96 * pred$se.fit
  df$upper <- pred$fit + 1.96 * pred$se.fit
  return(df)
}

preds2_ci <- preds2_filtered %>%
  group_by(Species) %>%
  do(compute_ci(.))

# Plot with manual confidence intervals and consistent legend
TPC <- ggplot(data = mydata_filtered) +
  geom_point(aes(x = K - 273.15, y = rate, color = Species), size = 2) +  # Points colored by Species
  geom_line(data = preds2_ci, aes(x = K - 273.15, y = fit, color = Species), size = 1) +
  geom_ribbon(data = preds2_ci, aes(x = K - 273.15, ymin = lower, ymax = upper, fill = Species), alpha = 0.2, color = NA) +
  scale_color_manual(values = color_palette, labels = custom_labels) +
  scale_fill_manual(values = color_palette, labels = custom_labels) +
  theme_bw(base_size = 12, base_family = 'Helvetica') +
  ylab(expression(plain("Rate (log " * mu * "mol " * g^-1 * " hr"^-1 * ")"))) +  # Custom y-axis label
  xlab('Temperature (ºC)') +
  theme(strip.text.x = element_blank(),
        strip.background = element_blank(),
        legend.position = "right",  # Adjusted to show legend
        legend.title = element_text(size = 16),  # Adjust legend title size
        legend.text = element_text(size = 14, hjust = 0),  # Left-align legend text
        legend.justification = c(0.4, 0.4),  # Align legend box to the left
        legend.box.margin = margin(0, 0, 0, 0),
        axis.text.x = element_text(face = "plain", color = "black", size = 18), 
        axis.text.y = element_text(face = "plain", color = "black", size = 18),
        axis.title.x = element_text(color = "black", size = 23, face = "plain"),
        axis.title.y = element_text(color = "black", size = 23, face = "plain"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black")) +
  scale_y_continuous(limits = c(0, 4.1), breaks = seq(0, 4, by = 1)) +
  labs(col = "Species", fill = "Species")

print(TPC)
```

# Removing outliers
```{r}
#Outlier Removal

topt_Tpla_pred<- c(36.85329013, 17.19091269, 36.54186585, 33.86951498, 34.40569888, 36.63455529, 34.42686951, 34.28095702, 37.28888324)
threshold <- 1.5
all_outliers <- c()
while (TRUE) {
  mean_data <- mean(topt_Tpla_pred)
  sd_data <- sd(topt_Tpla_pred)
  z_scores <- (topt_Tpla_pred - mean_data) / sd_data
 outliers <- which(abs(z_scores) > threshold)
if (length(outliers) == 0) {
    break
}
 cat("Outliers:", topt_Tpla_pred[outliers], "\n")
  all_outliers <- c(all_outliers, topt_Tpla_pred[outliers])
  topt_Tpla_pred <- topt_Tpla_pred[-outliers]
}

cat("All outliers:", all_outliers, "\n")

topt_Vmel_pred<- c(24.55483845, 38.84118587)
threshold <- 1.5
all_outliers <- c()
while (TRUE) {
  mean_data <- mean(topt_Vmel_pred)
  sd_data <- sd(topt_Vmel_pred)
  z_scores <- (topt_Vmel_pred - mean_data) / sd_data
 outliers <- which(abs(z_scores) > threshold)
if (length(outliers) == 0) {
    break
}
 cat("Outliers:", topt_Vmel_pred[outliers], "\n")
  all_outliers <- c(all_outliers, topt_Vmel_pred[outliers])
  topt_Vmel_pred <- topt_Vmel_pred[-outliers]
}

cat("All outliers:", all_outliers, "\n")

```


# Building your statistical models
```{r}
# Parameter modeling
library(dplyr)

#Make sure you are working with the correct dataset (i.e., no outliers)
Topt_data <- read.csv("Topt_nooutliers.csv")

Topt_data_predation <- Topt_data %>%
  filter(Experiment %in% c("Predation"))


# Convert Species to factor
Topt_data_predation$Species <- as.factor(Topt_data_predation$Species)

# Remove "Vmelones" and "Hcumingi" from the dataset
Topt_data_predation_filtered <- Topt_data_predation %>%
  filter(Species != "Vmelones" & Species != "Hcumingi")

# Check the unique levels of Species after filtering
levels(Topt_data_predation_filtered$Species)

# Now, Topt_data_predation_filtered contains the dataset with "Vmelones" and "Hcumingi" removed


# Relevel the factor variable
Topt_data_predation_filtered$Species <- relevel(Topt_data_predation_filtered$Species, ref = "Tplanospira")


# Fit the linear model
glm <- glm(Topt ~ Species, data = Topt_data_predation_filtered, family = Gamma(link = "log"))

R_lm <- lm(log(Topt) ~Species, data = Topt_data_predation_filtered)

R <- lm(Topt ~Species, data = Topt_data_predation_filtered)


AIC(glm, R_lm, R)

BIC(glm, R_lm, R)

# Obtain the residuals
residuals_log <- residuals(R_lm )
residuals_lm <- residuals(R)
#residuals_log <- residuals(R_log)

# Create a QQ plot using qqnorm() and qqline()

qqnorm(residuals_log, main = "QQ Plot for Log")
qqline(residuals_log, col = 2)

qqnorm(residuals_lm, main = "QQ Plot for LM")
qqline(residuals_lm, col = 2)
```


# Selecting your model
```{r}
# Selected model
R_lm <- lm(log(Topt) ~Species, data = Topt_data_predation_filtered)
```


# Statistical analysis on the chosen model
```{r}
#To compare the means of the levels of the Species variable
TukeyHSD(aov(R_lm))

#Provides an overall assessment of the significance of the model and the individual predictors
anova(R_lm)

#Gives you detailed information about the coefficients, standard errors, t-values, and p-values for each predictor
summary(R_lm)
```


# Pairwise comparisons for significance between groups
```{r}
#Pairwise comparisons between the levels of the Species variable to identify which groups are significantly different from each other

#install.packages("multcomp")
library(multcomp)

# Create a contrast matrix for all pairwise comparisons between species
cm <- glht(R_lm, linfct = mcp(Species = "Tukey"))

# Perform the Tukey HSD test
TukeyHSD <- summary(cm)

# View the results
print(TukeyHSD)
```
